---
title: "STAT 6620 HW#8"
author: "Lanqin Zhao,   NetID: ae8275"
output: html_notebook
---

### Identifying Frequently-Purchased Groceries Using Association Rules

## Step 1 - collecting data ----
Our market basket analysis will utilize the purchase data collected from one month of
operation at a real-world grocery store. The data contains 9,835 transactions or about
327 transactions per day.The dataset  was adapted from the Groceries dataset in
the arules R package.

## Step 2: Exploring and preparing the data ----
Explore and Prepare the data for modeling.

# Data preparation - creating a sparse matrix for transaction data

```{r}
library(arules)
groceries <- read.transactions("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml13/groceries.csv", sep = ",")
 summary(groceries)

```


```{r}
# look at the first five transactions
inspect(groceries[1:5])
```

```{r}
# examine the frequency of items 
itemFrequency(groceries[, 1:3])
```
# Visualizing item support - item frequency plots

```{r}
# plot the frequency of items
itemFrequencyPlot(groceries, support = 0.1)
itemFrequencyPlot(groceries, topN = 20)
```

#Visualizing the transaction data - plotting the sparse matrix

```{r}
# a visualization of the sparse matrix for the first five transactions
image(groceries[1:5])
```

```{r}
# visualization of a random sample of 100 transactions
image(sample(groceries, 100))
```

## Step 3: Training a model on the data ----
First, we attempted to use the default settings of support = 0.1 and confidence = 0.8, we  ended up with a set of zero rules. Then, we set support =0.006, confidence = 0.25, minlen = 2, get 463 rules.


```{r}
library(arules)

# default settings result in zero rules learned
apriori(groceries)
```

# set better support and confidence levels to learn more rules
```{r}
groceryrules <- apriori(groceries, parameter = list(support =
                          0.006, confidence = 0.25, minlen = 2))
groceryrules


```

## Step 4: Evaluating model performance ----
To obtain a high-level overview of the association rules, we can use summary().

```{r}

# summary of grocery association rules
summary(groceryrules)

# look at the first three rules
inspect(groceryrules[1:3])
```


## Step 5: Improving model performance ----
It's useful to be able to sort rules according to different criteria, and get
them out of R into a form that can be shared with marketing teams and examined in
more depth. In this way, we can improve the performance of our rules by making the
results more actionable.

# Sorting the set of association rules
```{r}
# sorting grocery rules by lift
inspect(sort(groceryrules, by = "lift")[1:5])

```
#Taking subsets of association rules

```{r}
# finding subsets of rules containing any berry items
berryrules <- subset(groceryrules, items %in% "berries")
inspect(berryrules)
```

#Saving association rules to a file or data frame
```{r}

# writing the rules to a CSV file
write(groceryrules, file = "groceryrules.csv",
      sep = ",", quote = TRUE, row.names = FALSE)

# converting the rule set to a data frame
groceryrules_df <- as(groceryrules, "data.frame")
str(groceryrules_df)
```





