---
title: "STAT 6620 HW#8"
author: "Lanqin Zhao,   NetID: ae8275"
output: html_notebook
---

### Perform Optical Character Recognition using SVM

## Step 1 - collecting data ----
We will utilize letterdata.csv donated to the UCI Machine Learning Data
Repository (http://archive.ics.uci.edu/ml) by W. Frey and D. J. Slate. The
dataset contains 20,000 examples of 26 English alphabet capital letters as printed
using 20 different randomly reshaped and distorted black and white fonts. There are 16 statistical attributes are recorded. 

## Step 2: Exploring and preparing the data ----
Prepare the data for modeling.

# Import data
```{r}
letters <- read.csv("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml11/letterdata.csv")
str(letters)
```

# divide into training and test data
```{r}
letters_train <- letters[1:16000, ]
letters_test  <- letters[16001:20000, ]

```

## Step 3: Training a model on the data ----
We used the simple linear kernel function first.

# begin by training a simple linear SVM

```{r}
library(kernlab)
letter_classifier <- ksvm(letter ~ ., data = letters_train,
                          kernel = "vanilladot")
# look at basic information about the model
letter_classifier

```

## Step 4: Evaluating model performance ----
The classifier correctly identified the letter in 3,357 out of the 4,000 test records, and the accuracy is about 84 percent.

# predictions on testing dataset
```{r}
letter_predictions <- predict(letter_classifier, letters_test)

head(letter_predictions)

table(letter_predictions, letters_test$letter) #"confusion matrix"

```


# look only at agreement vs. non-agreement
# construct a vector of TRUE/FALSE indicating correct/incorrect predictions
```{r}
agreement <- letter_predictions == letters_test$letter
table(agreement)
prop.table(table(agreement))
```

## Step 5: Improving model performance ----
We use  a more complex kernel function: Gaussian RBF kernel. The new classifier correctly identified the letter in 3,725 out of the 4,000 test records, and the accuracy is about 93 percent. These are much better than those the previous model performed.

```{r}
set.seed(12345)
letter_classifier_rbf <- ksvm(letter ~ ., data = letters_train, kernel = "rbfdot")
letter_predictions_rbf <- predict(letter_classifier_rbf, letters_test)

agreement_rbf <- letter_predictions_rbf == letters_test$letter
table(agreement_rbf)
prop.table(table(agreement_rbf))


```


