---
title: "STAT 6620 HW#5"
author: "Lanqin Zhao,   NetID: ae8275"
output: html_notebook
---

### predicting medical expenses using linear regression

## Step 1 - collecting data ----
We will utilize insurance.csv file. It includes 1,338 examples of beneficiaries currently enrolled in the insurance plan, with features indicating characteristics of the patient as well as the total medical expenses charged to the plan for the calendar year. The features are:
. age: An integer indicating the age of the primary beneficiary (excluding those above 64 years, since they are generally covered by the government).
. sex: The policy holder's gender, either male or female.
. bmi: The body mass index (BMI), which provides a sense of how over- or under-weight a person is relative to their height. BMI is equal to weight (in kilograms) divided by height (in meters) squared. An ideal BMI is within the range of 18.5 to 24.9.
. children: An integer indicating the number of children/dependents covered by the insurance plan.
. smoker: A yes or no categorical variable that indicates whether the insured regularly smokes tobacco.
. region: The beneficiary's place of residence in the US, divided into four geographic regions: northeast, southeast, southwest, or northwest.

## Step 2: Exploring and preparing the data ----
Explore the data to see whether we can use the data to predict and prepare the data for modeling.
# Explore data
Import data

```{r}
#Import the CSV file.
insurance <- read.csv("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml10/insurance.csv", stringsAsFactors = TRUE)
str(insurance)

```
check for normality of dependent variable, insurance expenses. Because the mean value is greater than the median, this implies that the distribution of insurance expenses is right-skewed. The histogram  confirms this.

```{r}
# summarize the charges variable
summary(insurance$expenses)

# histogram of insurance charges
hist(insurance$expenses)

```

Exploring relationships among features - the correlation matrix
From the result, none of the correlations in the matrix are considered strong, but there are some
notable associations.
```{r}
cor(insurance[c("age", "bmi", "children", "expenses")])
```
```{r}
pairs(insurance[c("age", "bmi", "children", "expenses")])
```
```{r}
library(psych)
pairs.panels(insurance[c("age", "bmi", "children", "expenses")])
```

## Step 3: Training a model on the data ----
Using Linear regression to train a model on the data. The model shows that for each additional year of age, we would expect $256.80 higher medical expenses on average, assuming everything else is equal. Similarly, each additional child results in an average of $475.70 in additional medical expenses each year, and each unit
increase in BMI is associated with an average increase of $339.30 in yearly medical expenses, all else equal......
The results of the linear regression model make logical sense: old age, smoking, and obesity tend to be linked to additional health issues, while additional family member dependents may result in an increase in physician visits and preventive care such as vaccinations and yearly physical exams.
```{r}
ins_model <- lm(expenses ~ ., data = insurance)
# see the estimated beta coefficients
ins_model
```

## Step 4: Evaluating model performance ----
The model output provides three key ways to evaluate the performance of model:
1. The residuals:  the maximum error of 29981.7 suggests that the model under-predicted expenses by nearly $30,000 for at least one observation. On the other hand, 50 percent of errors fall within the 1Q and 3Q values, so the majority of predictions were between $2,850.90 over the true value and $1,383.90 under
the true value.
2. Estimated regression coefficient: model has several highly significant variables, and they seem to be related to the outcome in logical ways.
3. R-squared: the value is 0.7494, the model explains nearly75 percent of the variation in the dependent variable
Given the preceding three performance indicators, our model is performing fairly well.

```{r}
# see more detail about the estimated beta coefficients
summary(ins_model)

```

## Step 5: Improving model performance ----
There are three possible methods to improve model performance: adding non-linearrelationships,converting a numeric variable to a binary indicator,adding interaction effects.We try these by:
a.Added a non-linear term for age
b. Created an indicator for obesity
c. Specified an interaction between obesity and smoking
Relative to our first model, the R-squared value has improved from 0.75 to about 0.87. Similarly, the adjusted R-squared value,which takes into account the fact that the model grew in complexity, also improved
from 0.75 to 0.87. Our model is now explaining 87 percent of the variation in medical costs.
```{r}
# add a higher-order "age" term
insurance$age2 <- insurance$age^2

# add an indicator for BMI >= 30
insurance$bmi30 <- ifelse(insurance$bmi >= 30, 1, 0)

# create final model
ins_model2 <- lm(expenses ~ age + age2 + children + bmi + sex +
                   bmi30*smoker + region, data = insurance)

summary(ins_model2)

```

## R code from the Chapter Lab Sections 3.6.2 and 3.6.3
# 3.6.2
```{r}
#library(MASS)
#fix(Boston)
names(Boston )
#attach (Boston )
#lm.fit =lm(medv ~ lstat)
lm.fit =lm(medv~lstat ,data=Boston )
lm.fit
names(lm.fit )
confint (lm.fit)
predict(lm.fit, data.frame(lstat=(c(5 ,10 ,15) )),interval="confidence")
plot(lstat ,medv)
abline (lm.fit)
abline (lm.fit ,lwd =3)
abline (lm.fit ,lwd =3, col ="red ")
plot(lstat ,medv ,col ="red ")
plot(lstat ,medv ,pch =20)
plot(lstat ,medv ,pch ="+")
plot (1:20 ,1:20, pch =1:20)
par(mfrow =c(2,2))
plot(lm.fit)
plot(predict (lm.fit), residuals (lm.fit))
plot(predict (lm.fit), rstudent (lm.fit))
plot(hatvalues (lm.fit ))
which.max (hatvalues (lm.fit))     
     
```
# 3.6.3 Multiple Linear Regression

```{r}
lm.fit =lm(medv ~ lstat+age , data=Boston )
summary (lm.fit)
lm.fit =lm(medv ~ .,data=Boston )
summary (lm.fit)
library (car)
vif(lm.fit)
lm.fit1=lm(medv ~ . -age ,data=Boston )
summary (lm.fit1)
lm.fit1=update (lm.fit , ~ .-age)

```






