---
title: "STAT 6620 HW#7"
author: "Lanqin Zhao,   NetID: ae8275"
output: html_notebook
---

### Prodict the Strength of Concrete  using ANN

## Step 1 - collecting data ----
We will utilize concrete.csv file. It includes 1030 examples and 9 variables. The variables are:
cement      : num  141 169 250 266 155 ...
slag        : num  212 42.2 0 114 183.4 ...
ash         : num  0 124.3 95.7 0 0 ...
water       : num  204 158 187 228 193 ...
superplastic: num  0 10.8 5.5 0 9.1 0 0 6.4 0 9 ...
coarseagg   : num  972 1081 957 932 1047 ...
fineagg     : num  748 796 861 670 697 ...
age         : int  28 14 28 28 28 90 7 56 28 28 ...
strength    : num  29.9 23.5 29.2 45.9 18.3 ...

## Step 2: Exploring and preparing the data ----
Explore the data to see whether we can use the data to predict the strength and prepare the data for modeling.

# Import data
```{r}
concrete<- read.csv("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml11/concrete.csv")
# examine the launch data
str(concrete)
```
# custom normalization function and apply it 
```{r}
normalize <- function(x) { 
  return((x - min(x)) / (max(x) - min(x)))
}
# apply normalization to entire data frame
concrete_norm <- as.data.frame(lapply(concrete, normalize))

# confirm that the range is now between zero and one
summary(concrete_norm$strength)

# compared to the original minimum and maximum
summary(concrete$strength)

```
# create training and test data
```{r}
concrete_train <- concrete_norm[1:773, ]
concrete_test <- concrete_norm[774:1030, ]
```

## Step 3: Training a model on the data ----
We'll begin by training the simplest multilayer feedforward network with only a single hidden node.

```{r}
# train the neuralnet model
library(neuralnet)
set.seed(12345) # to guarantee repeatable results
concrete_model <- neuralnet(formula = strength ~ cement + slag +
                              ash + water + superplastic + 
                              coarseagg + fineagg + age,
                              data = concrete_train)


```

# visualize the network topology
```{r}
plot(concrete_model)

```

# Reference: http://www.r-bloggers.com/neuralnettools-1-0-0-now-on-cran/
```{r}
# alternative plot
library(NeuralNetTools)
# plotnet
par(mar = numeric(4), family = 'serif')
plotnet(concrete_model, alpha = 0.6)
```

## Step 4: Evaluating model performance ----
We use correlation to evaluate the model performance. The correlation is 0.806, which implies that our model is doing a fairly good job, even with only a single hidden node.

```{r}
# obtain model results
model_results <- compute(concrete_model, concrete_test[1:8 ])
# obtain predicted strength values
predicted_strength <- model_results$net.result
# examine the correlation between predicted and actual values
cor(predicted_strength, concrete_test$strength) 
```

# produce actual predictions by 

```{r}
head(predicted_strength)

concrete_train_original_strength <- concrete[1:773,"strength"]

strength_min <- min(concrete_train_original_strength)
strength_max <- max(concrete_train_original_strength)

head(concrete_train_original_strength)

# custom normalization function
unnormalize <- function(x, min, max) { 
  return( (max - min)*x + min )
}

strength_pred <- unnormalize(predicted_strength, strength_min, strength_max)
strength_pred
```

## Step 5: Improving model performance ----

We use a more complex ANN to improve performance. First, We increase the number of hidden nodes to five. The correlation is 0.924, which is much higher than 0.806, that of the previous model.
Then, we keep the network topology, and change the activation function, the correlation is 0.574, which is much worse than 0.806.

# a more complex neural network topology with 5 hidden neurons
```{r}

set.seed(12345) # to guarantee repeatable results
concrete_model2 <- neuralnet(strength ~ cement + slag +
                               ash + water + superplastic + 
                               coarseagg + fineagg + age,
                               data = concrete_train, hidden = 5, act.fct = "logistic")

# plot the network
plot(concrete_model2)


```
```{r}
# plotnet
par(mar = numeric(4), family = 'serif')
plotnet(concrete_model2, alpha = 0.6)

```

```{r}
# evaluate the results as we did before
model_results2 <- compute(concrete_model2, concrete_test[1:8])
predicted_strength2 <- model_results2$net.result
cor(predicted_strength2, concrete_test$strength)  

```

# try different activation function
```{r}

# a more complex neural network topology with 5 hidden neurons and a different function
set.seed(12345) # to guarantee repeatable results
concrete_model2 <- neuralnet(strength ~ cement + slag +
                               ash + water + superplastic + 
                               coarseagg + fineagg + age,
                             data = concrete_train, hidden = 5, act.fct = "tanh")

# evaluate the results as we did before
model_results2 <- compute(concrete_model2, concrete_test[1:8])
predicted_strength2 <- model_results2$net.result
cor(predicted_strength2, concrete_test$strength)  

```



