---
title: "STAT 6620 HW#8"
author: "Lanqin Zhao,   NetID: ae8275"
output: html_notebook
---

### finding teen market segments using k-means clustering

## Step 1 - collecting data ----
we will use a dataset representing a random sample of 30,000 U.S.high school students who had profiles on a well-known SNS in 2006.The data was sampled evenly across four high school graduation years (2006 through 2009) representing the senior, junior, sophomore, and freshman classes at the time of data collection.

## Step 2: Exploring and preparing the data ----
Explore and Prepare the data for modeling.

# import data
```{r}
teens <- read.csv("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml12/snsdata.csv")
str(teens)

```
# look at missing data 

```{r}
# look at missing data for gender variable
table(teens$gender)
table(teens$gender, useNA = "ifany")
```

```{r}
# look at missing data for age variable
summary(teens$age)
```

```{r}
# eliminate age outliers
teens$age <- ifelse(teens$age >= 13 & teens$age < 20,
                     teens$age, NA)

summary(teens$age)

```
# Data preparation - dummy coding missing values

```{r}
# reassign missing gender values to "unknown"
teens$female <- ifelse(teens$gender == "F" &
                         !is.na(teens$gender), 1, 0)
teens$no_gender <- ifelse(is.na(teens$gender), 1, 0)
# check our recoding work
table(teens$gender, useNA = "ifany")
table(teens$female, useNA = "ifany")
table(teens$no_gender, useNA = "ifany")

```

# Data preparation - imputing the missing values

```{r}
# finding the mean age by cohort
mean(teens$age) # doesn't work
mean(teens$age, na.rm = TRUE) # works

# age by cohort
aggregate(data = teens, age ~ gradyear, mean, na.rm = TRUE)

# create a vector with the average age for each gradyear, repeated by person
ave_age <- ave(teens$age, teens$gradyear,
                 FUN = function(x) mean(x, na.rm = TRUE))


teens$age <- ifelse(is.na(teens$age), ave_age, teens$age)

# check the summary results to ensure missing values are eliminated
summary(teens$age)

```
# Data preparation - preparing data for modeling
Only the 36 features that represent the number of times various interests appeared on the teen SNS profiles. For convenience, let's make a data frame containing only these features

```{r}
interests <- teens[5:40]
interests_z <- as.data.frame(lapply(interests, scale))
```

## Step 3: Training a model on the data ----
Start K=5.
```{r}
set.seed(2345)
teen_clusters <- kmeans(interests_z, 5)

```

## Step 4: Evaluating model performance ----
One of the most basic ways to evaluate the utility of a set of clusters is to examine
the number of examples falling in each of the groups. If the groups are too large or
too small, they are not likely to be very useful.For a more in-depth look at the clusters, we can examine the coordinates of the cluster centroids. From the results, the five clusters provided useful and novel insights into the interests of teenagers. By that measure, the algorithm appears to be performing quite
well.

```{r}

# look at the size of the clusters
teen_clusters$size
# look at the cluster centers
teen_clusters$centers
```

## Step 5: Improving model performance ----
We  now focus our effort on turning these insights our model produced into action. We examine how the cluster assignment relates to individual characteristics.

```{r}
# apply the cluster IDs to the original data frame
teens$cluster <- teen_clusters$cluster

# look at the first five records
teens[1:5, c("cluster", "gender", "age", "friends")]


```
# mean age by cluster
The mean age does not vary much by cluster, which is not too surprising as these teen identities are often determined before high school.

```{r}
aggregate(data = teens, age ~ cluster, mean)

```

# proportion of females by cluster
Recall that overall about 74 percent of the SNS users are female. Cluster 1, the so-called Princesses, is nearly 84 percent female, while Cluster 2 and Cluster 5 are only about 70 percent female. These disparities imply that there are differences in the interests that teen boys and girls discuss on their social networking pages.

```{r}
aggregate(data = teens, female ~ cluster, mean)

```
# mean number of friends by cluster
On an average, Princesses have the most friends (41.4), followed by Athletes (37.2) and Brains (32.6). On the low end are Criminals (30.5) and Basket Cases (27.7).

```{r}
aggregate(data = teens, friends ~ cluster, mean)

```




