---
title: "STAT 6620 HW#6"
author: "Lanqin Zhao,   NetID: ae8275"
output: html_notebook
---

### Identifying loan default using Logistic regression

## Step 1 - collecting data ----
We will utilize credit.csv file. It includes 1000 examples and 17  variables. The response variable is default.

## Step 2: Exploring and preparing the data ----
Explore the data to see whether we can use the data to identify loan default and prepare the data for modeling.

# Import data
```{r}
credit <- read.csv("http://www.sci.csueastbay.edu/~esuess/classes/Statistics_6620/Presentations/ml7/credit.csv")


# examine the credit data
str(credit)

```
# Prepare response variable
recode default variable: "No" into 0 and "Yes" to 1.

```{r}
# Fix the default variable to be 0 or 1

credit$default = as.numeric(credit$default)
credit$default = credit$default - 1
```
# Check if there are any missing values

```{r}

library(Amelia)
missmap(credit, main = "Missing values vs observed")

```
```{r}
# number of missing values in each column

sapply(credit,function(x) sum(is.na(x)))

# number of unique values in each column

sapply(credit, function(x) length(unique(x)))

```

# Prepare trainning and test data sets
```{r}
indx = sample(1:nrow(credit), as.integer(0.9*nrow(credit)))

credit_train = credit[indx,]
credit_test = credit[-indx,]

credit_train_labels = credit[indx,17]
credit_test_labels = credit[-indx,17]   

```

## Step 3: Training a model on the data ----

```{r}
# fit the logistic regression model, with all predictor variables

model <- glm(default ~.,family=binomial(link='logit'),data=credit_train)
model

summary(model)

anova(model, test="Chisq")

```

# drop the insignificant predictors, alpha = 0.10
```{r}


model <- glm(default ~ checking_balance + months_loan_duration + credit_history +  percent_of_income + age,family=binomial(link='logit'),data=credit_train)
model

summary(model)

anova(model, test="Chisq")
 
```

## Step 4: Evaluating model performance ----
Use confusion matrix, accuracy  and ROC to evaluate model performance. The accuracy is 75%, which is good. 

```{r}
# predictions on test data
fitted.results <- predict(model,newdata=credit_test,type='response')
fitted.results <- ifelse(fitted.results > 0.5,1,0)

# Confusion Matrix
library(gmodels)
CrossTable(credit_test$default, fitted.results ,
           prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,
           dnn = c('actual default', 'predicted default'))

# compute accuracy
misClasificError <- mean(fitted.results != credit_test$default)
print(paste('Accuracy',1-misClasificError))

```
```{r}
# ROC

library(ROCR)
pr <- prediction(fitted.results , credit_test_labels)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf)

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc
```

## Step 5: Improving model performance ----
The possible way to improve model performance might be getting more of the number of examples and feature variables.







